Finding ARG1 of Partitive Nouns in NomBank with DistilBERT

### 2022-05-04 src/paper.tex
Update

The baseline we use is a maximum entropy machine learning method in a token-by-token classification manner. Maximum entropy is a general technique for estimating probability distributions from data. Labeled training data
is used to derive a set of constraints for the model that
characterize the class-specific expectations for the distribution. To perform well, the model relies on the good representation of the features we feed it. In this paper, we select features including the word stem, the neighboring 2 words, POS tags, NG-BIO tags, whether capitalized and the position of the word in the sentence.

@@

The output of the question-answering model is the location (start index and end index) of the predicted answer. By locating the token on the original input token list with the start index and end index, we can reconstruct the ARG1 label of the sentence.

@@

>>>>

The baseline we use is a maximum entropy machine learning method in a token-by-token classification manner. Maximum entropy is a general technique for estimating probability distributions from data. Labeled training data
is used to derive a set of constraints for the model that
characterize the class-specific expectations for the distribution. To perform well, the model relies on the good representation of the features we feed it. In this paper, we select features including the word stem, the neighboring 2 words, POS tags, NG-BIO tags, whether capitalized and the position of the word in the sentence. Figure \ref{fig:maxent-feature-sample} shows the features we selected for the maximum entropy baseline.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{assets/maxent-feature-sample.png}
  \caption{Sample sentence features for maxent system}
  \label{fig:maxent-feature-sample}
\end{figure}

@@

The output of the question-answering model is the location (start index and end index) of the predicted answer. By locating the token on the original input token list with the start index and end index, we can reconstruct the ARG1 label of the sentence. The answer relocation is a bit more complicated due to the tokenization discrepancies between the original input and the tokenized input. This is explained in detailed in section \ref{section:data-preprocessing}.

@@

\subsection{Data Preprocessing}
\label{section:data-preprocessing}

Even though for deep neural network input wouldn't need much feature engineering, we still need to tokenize the text and convert the text into numerical representation so that it can be served as the input of the model. For DistilBERT-based pipelines, the tokenizer provided by DistilBERT is used to tokenize the text. However, the dataset is already tokenized and the way DistilBERT tokenizes a sentence could be different from how the sentences are already tokenized in the dataset. More specifically, the tokenizer from DistilBERT could tokenize the sentence into even smaller tokens than the tokenizer used in the dataset. Therefore, this could cause the output label of the model fail to align with the input given. Fortunately, the results of the DistilBERT tokenizer include the align information for the tokenization. With this information, we can reconstruct the tokenized sentence from the DistilBERT tokenizer output and match the model output with the original input.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{assets/dataset-one-qa-sample.png}
  \caption{One sample sentence for QA task}
  \label{fig:dataset-one-qa-sample}
\end{figure}

A special treatment for the question-answering pipeline is that the ARG1 label has to be transformed into the answer location (the start and end indexes). And the question prompt for every sample is the same "What is the ARG1 of the sentence?". The same question prompt is used for validation and testing. Figure \ref{fig:dataset-one-qa-sample} is one processed sample sentence for the QA task, inputs are padded in batched and stacked as matrix to be able to be processed by the model in batches. \verb|input_ids| are the numerized tokens of both the question prompt and the input text concatenated by \verb|102|. \verb|attention_mask| is the attention the model should be given to the tokens, \verb|0| means the token should not be considered by the model. \verb|start_position| and \verb|end_position| are the start and end indexes of the answer in the input text.

\subsection{Experiment Settings}

All deep models used in this paper are implemented in PyTorch and with HuggingFace transformers libraries. The pre-trained models are fine-tuned on a single NVIDIA P100 (16GB) GPU on the Google Colab platform. The deep models are fine-tuned for 10 epochs with the learning rate of \verb|2e-5| and weight decay of \verb|0.01|. For token classification tasks, the batch size is 64. For question-answering tasks, the input would be bigger with the question prompt, so the batch size is set to 32 to be able to fit into the GPU's RAM.
